API design for machine learning software: experiences from the scikit-learn project

### 2013-10-18 paper.tex
Copy edit core API.

Estimator initialization and actual learning are strictly separated,
in a way that is similar to partial function application:
an estimator is initialized from a set of named hyper-parameter values
(e.g., the $C$ constant in Support Vector Machines (SVM))
and can be considered a function
that maps these values to actual learning algorithms.
The constructor does not see any actual data.
All it does is attach the given parameters to the object.
For the sake of model inspection, hyper-parameters are set as public attributes,
which is especially important in model selection settings.
Default hyper-parameter values are provided for all built-in estimators.
These defaults are set to be relevant in many common
situations in order to make estimators effective \textit{out-of-box}.

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
It always returns the estimator object it was called on,
which now serves as a model of its input and can be used to perform predictions or transformations of input data.

@@

To illustrate the initialize-fit sequence,
let us consider a supervised learning task using logistic regression (LR).
Given the API defined above, solving this problem is as simple as the following
example.
\begin{pythoncode}
from sklearn.linear_model import LogisticRegression
clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"} for
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C},
the strength of the regularization) are not explicitly given and
thus set to the default values. Upon calling \texttt{fit}, a model is
learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
and stored on the object for later use.
%% (In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name);
to build a random forest on
the same data, one would simply replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
algorithms are all provided as transformers.  In our example,
to standardize the input \texttt{X\_train} to zero mean and unit variance
before fitting the LR estimator, one would write:
\begin{small}
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler
scaler = StandardScaler()
scaler.fit(X_train)
X_train = scaler.transform(X_train)
\end{pythoncode}
\end{small}
Of course, in practice, it is important to apply the same preprocessing to the
test data \texttt{X\_test}. Since a \texttt{StandardScaler} estimator stores the
mean and standard deviation that it computed for the training set, transforming
an unseen test set \texttt{X\_test} maps it into the appropriate region of
feature space:
\begin{pythoncode}
X_test = scaler.transform(X_test)
\end{pythoncode}
Transformers also include a variety of learning algorithms, such as
dimensionality reduction (PCA, manifold learning), kernel approximation,
and other mappings from one feature space to another.

>>>>

Estimator initialization and actual learning are strictly separated,
in a way that is similar to partial function application:
an estimator is initialized from a set of named hyper-parameter values
(e.g., the $C$ constant in Support Vector Machines, SVM)
and can be considered a function
that maps these values to actual learning algorithms.
The constructor does not see any actual data.
All it does is attach the given parameters to the object.
For the sake of model inspection, hyper-parameters are set as public attributes,
which is especially important in model selection settings.
Default hyper-parameter values are provided for all built-in estimators.
These defaults are set to be relevant in many common
situations in order to make estimators effective \textit{out-of-box}.

Actual learning is performed by the \texttt{fit} method. This method is called
with training data (e.g., supplied as two arrays \texttt{X\_train} and
\texttt{y\_train} in supervised learning estimators). Its task is to run a
learning algorithm and to determine model-specific parameters from the training
data and set these as attributes on the estimator object. As a convention, the
parameters learned by an estimator are exposed as public attributes with names
suffixed with a trailing underscore (e.g., \texttt{coef\_} for the
learned coefficients of a linear model),
again to facilitate model inspection.
In the partial application view,
\texttt{fit} is a function from data to a model of that data.
It always returns the estimator object it was called on,
which now serves as a model and can be used to perform predictions
or transformations of new data.

@@

To illustrate the initialize-fit sequence,
let us consider a supervised learning task using logistic regression.
Given the API defined above, solving this problem is as simple as the following
example.
\begin{footnotesize}
\begin{pythoncode}
from sklearn.linear_model import LogisticRegression
clf = LogisticRegression(penalty="l1")
clf.fit(X_train, y_train)
\end{pythoncode}
\end{footnotesize}
In this snippet, a \texttt{LogisticRegression} estimator is first initialized by
setting the \texttt{penalty} hyper-parameter to \texttt{"l1"} for
$\ell_1$ regularization. Other hyper-parameters (such as \texttt{C},
the strength of the regularization) are not explicitly given and
thus set to the default values. Upon calling \texttt{fit}, a model is
learned from the training arrays \texttt{X\_train} and \texttt{y\_train},
and stored on the object for later use.
%% (In this example, \texttt{fit} is called solely for its side effects.)
Since all estimators share the same interface, using a different learning algorithm is
as simple as replacing the constructor (the class name).
To build a random forest on
the same data, one would simply replace
\texttt{LogisticRegression(penalty="l1")} in the snippet above by
\texttt{RandomForestClassifier()}.

@@

Since it is common to modify or filter data before feeding it to a learning
algorithm, some estimators in the library implement a \textit{transformer}
interface which defines a \texttt{transform} method. It takes as input some new
data \texttt{X\_test} and yields as output a transformed version of
\texttt{X\_test}. Preprocessing, feature selection, feature extraction and dimensionality reduction
algorithms are all provided as transformers.  In our example,
to standardize the input \texttt{X\_train} to zero mean and unit variance
before fitting the logistic regression estimator, one would write:
\begin{small}
\begin{pythoncode}
from sklearn.preprocessing import StandardScaler
scaler = StandardScaler()
scaler.fit(X_train)
X_train = scaler.transform(X_train)
\end{pythoncode}
\end{small}
Of course, in practice, it is important to apply the same preprocessing to the
test data \texttt{X\_test}. Since a \texttt{StandardScaler} estimator stores the
mean and standard deviation that it computed for the training set, transforming
an unseen test set \texttt{X\_test} maps it into the appropriate region of
feature space:
\begin{pythoncode}
X_test = scaler.transform(X_test)
\end{pythoncode}
Transformers also include a variety of learning algorithms, such as
dimensionality reduction (PCA, manifold learning), kernel approximation,
and other mappings from one feature space to another.
